Approximating the Optimal State Abstraction

### 2016-04-30 icml_workshop/introduction.tex
Added MDP background and consistency with 3 desiderata

% Planning and RL is hard.
For this reason abstraction is of considerable interest in planning and \ac{RL} work~\yrcite{li2006towards,abelhershko2016approx,konidaris2007building,hauskrecht1998hierarchical,parr1998reinforcement,sutton1999between,konidaris2015constructing,Jiang2015,andre2002state}. Although planning for optimal behavior in \acp{MDP} is known to be P-Complete in the size of the state and action space~\cite{papadimitriou1987complexity,littman1995complexity}, and many \ac{RL} algorithms for solving \acp{MDP} are known to require a number of samples only polynomial in the size of the state space~\cite{strehl2009reinforcement}, many state spaces are known to grown super-polynomially with the number of variables that characterize the domain \cite{abel2015goal} and robust action spaces are often large or even continuous~\cite{antos2008fitted,konidaris2014constructing}. Abstraction, then, promises an alluring avenue through which planning and \ac{RL} techniques may gain traction in complex domains.

@@

% We propose desiderata.
In this work, we characterize what it means for an abstraction to be effective in planning and \ac{RL}. In particular, we propose  two desiderata that a theory of abstraction should satisfy. The first we call \textsc{Compressibility}, which stipulates that a theory on abstraction ought to achieve {\it compression}. The second we coin \textsc{Optimality}, which states that abstractions ought to maintain at least some notion of optimality.

% Outline: We summarize framework, propose extension and then ground out desiderata in framework.
Additionally, we summarize our previously proposed formal framework for state-abstraction methods, and demonstrate that the framework allows for careful formal analysis of temporally-extended abstractions. In particular the framework allows us to characterize what types of temporal abstraction do and do not satisfy the \textsc{Compressibility} and \textsc{Optimality}. Of special interest is proving what the necessary or sufficient conditions are for an option to be useful in planning or learning. \dnote{come back to dis sentence}

>>>>

% Planning and RL is hard.
For this reason abstraction is of considerable interest in planning and \ac{RL} work~\yrcite{li2006towards,abelhershko2016approx,konidaris2007building,hauskrecht1998hierarchical,parr1998reinforcement,sutton1999between,konidaris2015constructing,Jiang2015,andre2002state}. Planning and \ac{RL} problems tackle solving for optimal behavior in \acp{MDP}.

A \ac{MDP} is a 5-tuple, $\langle \mathcal{S}, \mathcal{A}, \mathcal{T}, \mathcal{R}, \gamma \rangle$. $\mathcal{S}$ is a finite state space; $\mathcal{A}$ is a finite set of actions available to the agent; $\mathcal{T}$ denotes $\mathcal{T}(s' \mid s,a)$, the probability of an agent transitioning to state $s' \in \mathcal{S}$ after applying action $a \in \mathcal{A}$ in state $s \in \mathcal{S}$; $\mathcal{R}(s,a)$ denotes the reward received by the agent for executing action $a$ in state $s$; $\gamma \in [0, 1]$ is a discount factor that determines how much the agent prefers future rewards over immediate rewards. We assume without loss of generality that the range of all reward functions is normalized to $[0,1]$. The solution to an \ac{MDP} is called a policy, denoted $\pi: \mathcal{S} \mapsto \mathcal{A}$. The objective of an agent is to solve for the policy that maximizes its expected discounted reward from any state, denoted $\pi^*$.

 Although planning for optimal behavior in \acp{MDP} is known to be P-Complete in the size of the state and action space~\cite{papadimitriou1987complexity,littman1995complexity}, and many \ac{RL} algorithms for solving \acp{MDP} are known to require a number of samples only polynomial in the size of the state space~\cite{strehl2009reinforcement}, many state spaces are known to grown super-polynomially with the number of variables that characterize the domain \cite{abel2015goal} and robust action spaces are often large or even continuous~\cite{antos2008fitted,konidaris2014constructing}. Abstraction, then, promises an alluring avenue through which planning and \ac{RL} techniques may gain traction in complex domains.

@@

% We propose desiderata.
In this work, we characterize what it means for an abstraction to be effective in planning and \ac{RL}. In particular, we propose three desiderata that a theory of abstraction should satisfy. The first we call \textsc{Compressibility}, which stipulates that a theory on abstraction ought to achieve {\it compression}. The second we coin \textsc{Optimality}, which states that abstractions ought to maintain at least some notion of optimality. The last we term \textsc{Efficient-Computability} which states that an abstraction ought to be tractably computable.

% Outline: We summarize framework, propose extension and then ground out desiderata in framework.
Additionally, we summarize our previously proposed formal framework for state-abstraction methods, and demonstrate that the framework allows for careful formal analysis of temporally-extended abstractions. In particular the framework allows us to characterize what types of temporal abstraction do and do not satisfy \textsc{Compressibility}, \textsc{Optimality} and \textsc{Efficient-Computability}. Of special interest is proving what the necessary or sufficient conditions are for an option to be useful in planning or learning. \dnote{come back to dis sentence}
